
---
title: "Data Science for business - Final Project"
author: "Andrea Barabas (153651)"
date: "3 March 2016"
output: html_document
---
#Data Source and Description

In the framework of this project, I chose a publicly available dataset from [Kaggle](https://www.kaggle.com/c/airbnb-recruiting-new-user-bookings/data). I did not follow the rules of the competition, instead I focused on an exploratory data analysis, feature engineering and modelling based on the material the Data Science lectures covered.

New users on Airbnb can book a place to stay in 34,000+ cities across 190+ countries. By accurately predicting where a new user will book their first travel experience, Airbnb can share more personalized content with their community, decrease the average time to first booking, and better forecast demand.

The training set consists of 15 variables that help predict the country destination (+1) of all the new users with first activities after 7/1/2014. The Competition also contained three other datasets, web sessions log for users, summary statistics of destination countries as well as summary statistics of users' age group, gender, country of destination. 
The number of observations in the training set is 213451 whilst the test set contains 62096 observations and 15 variables.

*Predictor variables:
+ id: user id
+ date_account_created: the date of account creation
+ timestamp_first_active: timestamp of the first activity
+ date_first_booking: date of first booking
+ gender
+ age
+ signup_method
+ signup_flow: the page a user came to signup up from
+ language: international language preference
+ affiliate_channel: what kind of paid marketing
+ affiliate_provider: where the marketing is e.g. google, craigslist, other
+ first_affiliate_tracked: whats the first marketing the user interacted with before the signing up
+ signup_app
+ first_device_type
+ first_browser
+ country_destination: this is the target variable you are to predict

#Exploratory data analysis (EDA)

```{r, include=FALSE}
rm(list=ls())
library(plyr)
library(dplyr)
library(data.table)
library(readr)
library(dplyr)
library(reshape2)
library(ggplot2)
library(RColorBrewer)
```

##Loading datasets
```{r, echo=FALSE}
air_train<- read.csv('/Users/Andi/Desktop/airbnb_train.csv')
```
##Searching for missing values
```{r, echo=FALSE}
na_count <-sapply(air_train, function(y) sum(length(which(is.na(y)))))
na_count <- data.frame(na_count)
na_count
air_train$timestamp_first_active <- as.Date(strptime(air_train$timestamp_first_active,format="%Y%m%d%H%M%S"))
air_train$date_account_created <- as.Date(air_train$date_account_created)
```
##Predictor: Age
```{r}
length(which(air_train$age > 100))
length(which(air_train$age < 18))
air_train$age <- as.factor(air_train$age)
summary(air_train$age)
ggplot(air_train) + geom_bar(aes(x = age),fill="deepskyblue")
```
Due to the large number of missing values as well as the thousands of values well above 100 years, I decided to get rid of the age column.
###Heatmap for age and date of accounts created
```{r}
#taking out outliers
condition1 <- air_train[,6] >=18 & air_train[,6] <=100
air3 <- air_train[condition1,]
View(air3)

summary(air4)

# Heatmap on age and date of accounts dreated
air_train2 <- as.data.table(air3)
air_train2[, age:= as.numeric(age)]
air_train2$date_account_created <- as.Date(air_train2$date_account_created)
heatmapData <- as.matrix(table(air_train22$age,air_train$date_account_created)^.2)
p <- qplot(x=Var2, y=Var1, data=melt(heatmapData), fill=value, geom="tile", 
           xlab = "Date Account Created", ylab = "User Age", 
           main = "Number of Accounts Created Over Time Across User Ages")
p <- p + scale_x_discrete(breaks=levels(as.factor(air_train2$date_account_created))[c(TRUE, rep(FALSE, 90))], 
                          labels=levels(as.factor(air_train2$date_account_created))[c(TRUE, rep(FALSE, 90))])
p <- p + theme(axis.text.x  = element_text(angle=90, hjust=0.5, size=10))
p <- p + scale_fill_gradient(low="white", high="orangered")
print(p)
```

```{r, echo=FALSE}
air_train <- subset(air_train, select = -c(age))
```
#Predictor: signup_method
```{r, eval=FALSE}
air_train[,.N, by=signup_method] #basic, facebook, google

air_train %>% group_by(country_destination, `signup_method`) %>% 
  summarize(n = n()) %>% as.data.frame
```

```{r, eval=FALSE}
ggplot(air_train, aes(x=country_destination, fill=signup_method)) +
  geom_bar(position="dodge", colour="black") +
  scale_fill_brewer(palette="Pastel1")
```

```{r, eval=FALSE}
air_train2 <- air_train %>%
  filter(country_destination != "NDF")
air_train2 <-data.table(air_train2)

air_train2$date_first_booking <- as.Date(air_train2$date_first_booking)
```
##By taking out the NDF country destination:
```{r, eval=FALSE}
ggplot(air_train2, aes(x=country_destination, fill=signup_method)) +
  geom_bar(position="dodge", colour="black") +
  scale_fill_brewer(palette="Pastel1")
```

#Predictor: gender
```{r, eval=FALSE}
air_train2[,.N, by=gender] #-unknown-, FEMALE, MALE, OTHER

air_train2 %>% group_by(country_destination, `gender`) %>% 
  summarize(n = n()) %>% as.data.frame

air_train2 <- air_train2 %>%
  filter(gender != "-unknown-")

air_train2 <- air_train2 %>%
  filter(gender != "OTHER")
air_train2 %>% group_by(country_destination, `gender`) %>% 
  summarize(n = n()) %>% as.data.frame
```

```{r, eval=FALSE}
ggplot(air_train2, aes(x=country_destination, fill=gender)) +
  geom_bar(position="dodge", colour="black") +
  scale_fill_brewer(palette="rainbow(30)")
```

#Predictor: signup_app
```{r, eval=FALSE}
air_train2[,.N, by=signup_app]

air_train2 %>% group_by(country_destination, `signup_app`) %>% 
  summarize(n = n()) %>% as.data.frame
```

```{r, eval=FALSE}
ggplot(air_train2, aes(x=country_destination, fill=signup_app)) +
  geom_bar(position="dodge", colour="black") +
  scale_fill_brewer(palette="rainbow(30)")
```
##Predictor: language
```{r, eval=FALSE}
air_train2[,.N, by=language]
```

```{r, eval=FALSE}
ggplot(air_train2, aes(x=country_destination, fill=language)) +
  geom_bar(position="dodge", colour="black") +
  scale_fill_brewer(palette="rainbow(30)")
```

##Predictor: affiliate_provider

